Dynamic capabilities and competitive advantage

### 2019-10-21 article4_tommylars.tex
after tommy meeting first draft

0.343
In a recent "state of the art" literature review on dynamic capabilities
\cite{Schilke2018} state that much of the attention and interest surrounding the theory is
that it offer a route to competitive advantage which they argue is "the virtual Holy Grail
of strategic management" (p. 393). And indeed a considerable body of the extant literature
concerns itself with exactly this path to competitive advantage (refs from
shcilke). Still, a discourse regarding the contingencies of the dynamic
capabilities-competitive advantage (DC-CA) nexus (?) has panned out in the literature,
most notably coined "the elephant in the room" \citep{Peteraf2013}: Under which conditions
will dynamic capabilities lead to competitive advantage?

\cite{Peteraf2013} argue that the seminal papers in the dynamic capabilities theory
tradition (\cite{Teece1997} and \cite{Eisenhardt2000}) differ considerably in their view
of to what extent the DC-CA nexus exist and under what conditions it is functional. A key
tenant of this discourse is related to the level of routinization and the ability to
satisfy the VRIO condition \citep{Barney1991a} of the resource-based view
\citep{Peteraf2013}.

In this paper we aim to contribute to dissecting the elephant by asking how does the DC-CA
nexus differ between firms and what determines these differences?

We do this by exploring two paths from dynamic capabilities to competitive advantage: the
\emph{evolutionary path} and the \emph{market path}. The former is thoroughly described
and explored in the literature (refs) and stems from the core idea of the dynamic
capabilities theorizing. In this path dynamic capabilities work to change the operating
routines of the firm and hence contributing to it's competitive advantage. We argue that
this path is more suited for less dynamic environments because of its path dependent
functioning.

The \emph{market path} works not through shaping operational capabilities and routines,
but by enabling the firm to act in factor markets for new capabilities and
resources. Hence, it is less path dependent and can be more rapidly
deployed. Consequently, it is more prevalent in dynamic environments.

This paper contributes to the extant literature on the link between dynamic capabilities
and competitive advantage in three ways. First, provides an empirical test to the
drive-train model of \cite{DiStefano2014} on how dynamic capabilities can operate both
quickly and with high frequency more directly on the strategic direction of the firm by
acting in factor markets for important resources, and also more slowly through the
evolutionary path of changing operational capabilities. Second, by conceptually clarifying
the different paths to competitive advantage by linking it closely to the role of
capability evolution and the VRIO condition. Finally, we.........

CLARIFIES THE ROLE OF THE ENVIRONMENT AND MEDIATIONS HAPPENING

%% How does dynamic capabilities lead to competitive advantage?
%% Dynamic capabilities are proposed to confer a competitive advantage by adding unique value to the firm through systematic change, which enhances operational efficiency and enables increased alignment with the environment (Di Stefano et al., 2014; Peteraf et al., 2013)

%% Not all organizations possess them (Collis, 1996); their path dependencies, intangibility, complexity, and organizational specificity make them hard to imitate (Gibson & Birkinshaw, 2004; Helfat & Winter, 2011); and few other means allow organizations to purposefully change on a continuous basis (Day & Wensley, 1988; Helfat et al., 2007).

%% Consistent with the theoretical positions of Eisenhardt and Martin (2000), Helfat and Peteraf (2003), Zahra et al. (2006), and Zott (2003), among others, these researchers have argued that dynamic capabilities’ immediate purpose is to change the resource base, and that this change in the resource base, in turn, explains performance variations. According to this argument, resource changes serve as mediators through which dynamic capabilities affect performance. We

%% One of the major unresolved issues in dynamic capabilities theory is to what extent and
%% how competitive advantage is the ultimate outcome of dynamic capabilities. This very idea
%% has

%% This paper asks two questions: to what extent does a direct lnk between DC and CA exist,
%% and what determines variations in the role of mediation between them?

%% The idea that Dc impacts CA is uncontroversial (not). Little is known about how this
%% happens. The theory holds that this happens through xreating extending and modifying the
%% resource base. YEt, the change in firm capabilities can themselves lead to CA due to their
%% evlutionary nature.

%% Variations in the share of mediation stems from variation sin the environment. When envdyn
%% is high firms need to act faster and hence tends to choose strategies including resource
%% orchestration buying in factor markets cause it tkes more time to develop OR.

%% In less dynamism the role of OR increases and leads to a significant proportion of CA
%% stemming throhg the mediating effect of OR

%% There is ongoing interest in (dynamic) capabilities and how they influence firm
%% performance \citep{Fainshmidt2016a,Teece2014}. Indeed, \emph{dynamic capabilities} (DC),
%% and its front runner, the resource-based view \citep{Barney2001,Helfat2007}, is now
%% considered one of the major theories in strategic management
%% \citep{Schilke2018}. Arguably, the key reason is that this theory clearly articulates how
%% firms can alter their operating routines through dynamic capabilities. Thus, a key
%% ingredient in this theory is that DCs can alter firms operating routines, with subsequent
%% performance implications.

We demonstrate that two paths indeed do exist and that their relative importance varies
with the degree of environmental dynamism facing the firm.

@@

%% We can use the drive train to determine hwo DC works differently under different levels of
%% dynamism. It is not that DC does not work under stable conditions as EM claims. Is only
%% that the effect runs stringer through the evolutionary path. Whereas the market path is
%% stringer under higher dynamism

One contentious issue surrounding the DC theory is the extent to which differing views on
the nature of dynamic capabilities \citep{Peteraf2013} as well as their outcome. One
strand of contributions sees DC as routines that are simple in nature and more as best
practices to consider (EM). Thus, dynamic capabilities are more like behavioral patterns
and less idiosyncratic to the firm and will consequently not be able to satisfy the VRIO
condition for competitive advantage to emerge. Dynamic capabilities as such routines are
short lived, unstable and “substitutable” (EM: 1110) thus violating a key VRIN condition,
and they are more “more homogeneous (..) than is usually assumed” (EM: 1116). However, the
simplicity of simple rules also makes them faster do deploy and put in motion, albeit
potentially unstable. To the extent dynamic capabilities manifest themselves as simple
rules they cannot alone create competitive advantage, but could very well generate
strategic change by facilitating acquisition of resources in factor markets. One example
is the role of post-acquisition routines where routines to identify and deal with
acquisition partners are a source of generating strategic change (Eisenhardt and Sull
2001, ref). Another example is (see refs from Schilke 2018). The functioning of this
expression of dynamic capabilities are more latent and hence mainly observable ex
post. Hence we can see this as direct route to competitive advantage as this expression of
dynamic capabilities work directly in the factor market in capturing resources
needed for strategic change. We coin this direct connection as the \emph{market path}.

On the other hand, a stream of contributions see dynamic capabilities as evolutionary and
more complex routines (ref) or processes (ref) set to change underlying capabilities in
the face of changing conditions (TPS). This notion stems from evolutionary economics
\citep{Winter2003,Nelson1982} and the behavioral theory of the firm \citep{Cyert1963}
holding that path dependency and complex processes for changing underlying operating
routines are a source of competitive advantage for the firm. Along this indirect path
change is gradual and running through changes in underlying routines. We coin this the
\emph{evolutionary path}.

These paths from dynamic capabilities to competitive advantage are different expressions
of dynamic capabilities with different mediums of intermediation. The \emph{market path}
works in the market and his hence not typically altering lower order capabilities and
resources in the firm directly, but rather acquiring new ones. The source of competitive
advantage along this path stems from the interaction with the other, \emph{evolutionary
  path} in a dynamic bundle \citep{Peteraf2013}. Such interactions create a bundle of
capabilities and routines that are socially complex and hard to imitate
\citep[p. 320]{DiStefano2014}. Moreover, the interactions between the two paths of dynamic
capabilities is separable with respect to what intermediation exist. Consequently,
separate paths exist albeit a part of a dynamic bundle.

\cite{DiStefano2014} propose an integrated framework where simple routines and more
complex processes could contribute simultanously to the firms outcome, specifically its
competitive advantage. In this framework, coined the "organizational drivetrain" the
authors envision the movement of an organization as set in motion by a drivetrain system
not unlike that of a bicycle.

Building on the notion of an organizational drivetrain we argue that dynamic capabilities
can indeed impact the organizational movement or change through two distinct but
interconnected mechanisms (i.e. the crank set and the rear shifter). The two mechanisms
consitutting the drivetrain is consistent with our notion of \emph{evolutionary path} and the \emph{market path}. The
evolutionary path is instigated by more complex processes setting in motiong a chance in
the lower order (operational) routines \cite{Collis1994,Winter2003}. In the drivetrain
metaphor this path corresponds to change stemming from the front crank set of thich there
are fewer gears, but with larger impact. This path is slower
and yields a more idiosyncratic set of organizational routines which is very much in line
with the idea put forth in the seminal founding papers \citep{Teece1997,Winter2003}.

Seeing the bike rider and her efforts to handle the levers of control (i.e. the gear
shifter) as the dynamic capability, and consider it not only as a management effort, but
that of the organization as a whole. The process of changing front gears are more profound
(i.e. evolutionary) and costly, and it is for larger and slower strategic
changes. Operating with simple routines in factor markets, however, is quicker and the
rider does this on a more frequent basis with less effort.

H1: Dynamic capabilities works to create competitive advantage through two distinctly
different paths.

\subsection{The two paths of dynamic capabilities}
The objective of strategic change is to enhance a firm’s competitive advantage
(ref). Doing so over time makes for the evolutionary fitness of the firm
\citep{Helfat2007}. When firms are developing capabilities for strategic change through
creating, extending and modifying the resource base this means making decisions to develop
operational routines. Such routines can in turn stem from an inherent learning process
evolutionary distinct to the firm. Operating routines can thus lead to competitive
advantage because of their evolutionary character \citep{Nelson1982,Winter2003}. To the
extent operating routines stems from idiosyncratic evolutionary process of individual
firms rather than being the result of investments in factor markets, they can satisfy the
VRIO condition and thus be a source of competitive advantage. In such instances, dynamic
capabilities work through operating routines in creating competitive advantage. This
indirect impact on competitive advantage is very much at the core of the DC literature
(ref). Firms with high dynamic capabilities will be able to better change operating
routines to facilitate strategic change in response to changing conditions facing the
firm. Dynamic capabilities then creates VRIO operating routines and hence competitive
advantage. When this happens, firms changes their operating routines in a evolutionary
making operating routines themselves OR itself satisfies the VRIO condition.

\emph{H2: Dynamic capabilities has an indirect effect on competitive advantage through
  changing the operating routines of the firm.}

On the other hand, resources and capabilities can to a certain extent be captured in
factor markets through investment decisions (ref) or through alliances (ref) or M\&A
activates (ref). By acquiring resources through investments, the operating routines of the
firms may not be directly affected. When these acquisitions happen as a response to
systematic resource orchestration (ref), i.e. reshaping the resource mix to obtain
coordinated resource deployments \citep{Kor2005,Pan2006}, they would indeed have potential
performance implications for the firm. Moreover, when this process of resource acquisition
process of DC develops a way to acquire and mix resource for the purpose of adapting to
change, the resources themselves would not necessarily satisfy the VRIO condition. The way
in which this happens may be much harder to imitate due to the complexities such routines
exhibit. Although resources themselves exhibit clear equifinality in their impact on
competitive advantage, combined with the simple routines of the orchestration
(i.e. dynamic capabilities), the sum would likely satisfy the VRIO condition. This is
especially true if routines are deployed rapidly in changing market conditions. In other
words, there are several ways in which resources can be orchestrated to generate CA,
whereas improved OR will always mean improved competitiveness et ceteris paribus. In such
instances DC itself can lead to CA directly.

\emph{H3: Dynamic capabilities has a direct effect on competitive advantage through the
  process of resource orchestration and acquisition of resources in factor markets.}

In other words, we would expect DC to have both a direct and an indirect impact on CA. The
indirect impact stems from the \emph{evolutionary path} of complex processes, whereas the
direct impact stems from the \emph{market path}.  Although both the direct and the
indirect path stems from DC creating, extending and modifying the resource base (OR or
factor market resources), the difference in the path’s stems from to what extent the
mediating construct alone can by itself satisfy the VRIO condition. Operating routines
can, due to their evolutionary nature, in turn shaped by DC, achieve CA
\citep{Collis1994,Winter2003}. Factor market resources and market acquired capabilities
and best practices can not \citep{Eisenhardt2000,Peteraf2013}, but in conjuncture with the
routines of acting in factor markets as well as the speed of deployment, the sum of
dynamic capabilities and factor market action would lead to competitive advantage.

However, the interlinkages between simple routines (the market path) and the coplex
processes (the evolutionary path) would vary with the environment.

\subsection{Interplay under environmental dynamism}

\cite{DiStefano2014} note that "even if specific simple rules are unstable and ephemeral,
the system as a whole is not" and that the "real source of sustainable competitive ad- vantage for an enterprise is the difficulty of imitating and substituting for the entire dynamic bundle that the system represents" (p.sss).

But the interlinkages between the parts of the system (the evolutionary- and the market
path), are likely to vary with the environmental dynamism. Under moderatly dynamism
dynamic capabilities takes the form of best pracitces (e.g. aquisition targeting) and is
less important than under more dynamic environments. However, albeit best practices, the
way they are implemented in the factor markets are still idiosyncratic. And taken into
conjunction with the routinzed process of changing underlying routines, VRIO will very
likely emerge under more stable conditions. It is not that the market/direct expression
of dynamic capabilities are not present in stable environments, rather it is less pronounced and the indirect
mediated effect (evolutionary path) is more important. In other words, it is not neither
nor, but degrees of importance.

Moreover, in high velocity environments the role of the evolutionary path tends to
weaken as such path-dependent changes occur slower compared to the market path. From the
market path with its simple routines we can argue that creation of "new knowledge” to
“allow for emergent adaptation” (EM, p. 1116) is the consequence. Consequently, the market
path is stronger and more prevalent under higher levels of environmental dynamism.

\emph{H4: Dynamic capabilities works relatively stronger through the market path in highly
dynamic environments.}

>>>>

%% Extant literature on dynamic capabilities is battling with a fundamental conception and
%% prediction of the theory, namely if, when and how dynamic capabilities can lead to competitive
%% advantage.

The presence of a direct path from dynamic capabilities and competetive advantage is
somewhat puzzling. It eiter suggests ommited importrant mediating factors such as
resources, or that dynamic capabilities can act outside "creating, extending and modifying
the firm resources base". A growing consensus on the definition as well as the
theoretical mechanisms underpinning dynamic capabilities suggest that dc work indirectly
towards strategic change by changing firm resources. Hence, the presence of direct effects
seems increasingly odd. Still, seminal contributions to the field of dynamic capabilities have argued for both direct
\cite{Kor2005,Eriksson2014,Schilke2014a} and indirect (ref) effects of dynamic
capabilities on competitive advantage, although increasingly the bulk of the literature
focuses on the latter \citep{Schilke2018}.

 Resource orchestration is defined as "blablbalbla" (helfat)

This indirect link is typically understood as dynamic capabilities modifying, extending or
creating firm resources (helfat) (or lower-level operational capabilities (Winter,
collis)) and thus generating competitive advantage (Eriksson and others).The direct path,
however, is theoretically less explicitly dealt with in the extant literature although
empirical evidence keep suggesting that both direct (ref) and indirect (ref) links exist.

In this paper we explore to what extent a direct link between dynamic capabilities and
competitive advantage work, and under which conditions it is most pronounced. Specifically
the research question is:

{\bf How is a direct link between dynamic capabilities and competitive advantage working
  and under which conditions does it come into play or not?}

We argue that direct link exist as a response to the multilevel notion of strategy itself
and captures strategic change not typically associated with the resource based
view. Furthermore, we show that the prevalence of the direct link is a function of the
nature of underlying resources dynamic capabilities are aimed at "creating, extending and
modifying", specifically if they are valuable and idiosyncratic to the firm or not, and
the environment facing the firm.

Our findings indicate that the direct path is less prevalent in very dynamic environments
where the indirect path becomes far more important. Furthermore, the effect of dynamism is
partly contingent on the nature of the firm resources suggesting a more complex
understanding of how imitable and simple resources can play a role in generating
competetive advantage.

Our contributions are twofold. First, we explore different mediators by explicitly
modeling a causal chain with different types of underlying resources, as well as
under different environments. This is explicitly called for in the literature as "only
four percent of the articles in our sample explicitly examined possible causal mechanisms,
suggesting a significant gap in the literature" (Schilke 2018). This work sheds light on
the contingencies and boundary conditions of the link between dynamic capabilities and the
"holy grail of strategy" (Schilke et al), competetive advantage.

Second, we explore a theoretical explanation for why a direct link between dynamic
capabiliites and competetive advantage may exist. By leaning on empirical evidence
suggesting a multitude of explanations for performance differentials between firms, we
suggest that dyanmic capabilities can act outside the theorized path of "resource
orchestration" through strateic efforts such as industry positioning (porter) or
institutional leverage (Peng).

%%%%%%%%%%%%%%5

%%%%%%%%%%%%%%

%% This warrant the question of \emph{when}
%% direct and indirect links between dynamic capabilities and competitive advantage is most
%% pronounced.

%% Moreover, firm resources or capabiltiies themselves are far from a homogenous set and
%% varies among other things with respect to their uniqueness, inimatbility and value
%% (barney). Important contributions have thus argued for a deeper insighi into these
%% intermediaries and their mechanism calling for a "deeper insight into the variety of mechanisms that
%% underlie the performance effects of capabilities.” (Schilke, 2014a: 199). 
%% as well as \emph{what} different intermediaries exist.

%% This paper investigates the boundary conditions of the VRIO condition and how dynamic
%% capabilities and resources interact to create firm competitive advantage. We ask the
%% following research question:
%% Under which conditions are dynamic capabilities able to yield competietive advantage
%% through resource orchestration and what are the different outcomes of different types of resources.

%% We find thatthe dynamism surrounding hte firm represent an important boundary condition as
%% suggested in parts of the literature. However, DC consinutes to generate competetive
%% advnatage also in high-velocity environments, but doing increasingly so indirectly through
%% resource orchestration. Furthermore, we find that non-vrio resources are more likely to
%% yield competetive power in high velocity environments, and that vrio-resources tend to
%% maintain their CA under any condition.

%% We demonstrate that two paths indeed do exist and that their relative importance varies
%% with the degree of environmental dynamism facing the firm.

%% Under what cionditions are the direct path and indirect path most prevalant? Under which condi

%% The direct

@@

%{\bf How is a direct link between dynamic capabiliites and competetive advantage working
%  and under which conditions does it come into play or not?}

In the face of a growing consensus on the definition (ref) and outcome (ref) of dynamic
capabilities as working through mediators, the presence of direct effects argued by others
(ref) seems puzzling. However, the broader empirical literature on the antecedents of firm
performance offers an interesting point of departure. This literature concerns itself with
the relative importance of factors at the firm-level on the one hand, and those at the
industry level at the other.

The debate over the relative role of different levels of factors in accounting for the
major share of the variance in firms performance has been of particular importance in the
field of strategic management, most notably in the Strategic Management Journal
(M. A. Fitza 2014; Hansen and Wernerfelt 1989; McGahan and Porter 1997; Rumelt
1991). Overall, the accumulated empirical evidence show that firm factors matter far more
for firm performance than industry factors (McGahan and Porter 2002). These stylized facts
have been used as empirical support for the importance of idiosyncratic firm resource and
capabilities; leading up too recent formulation of the resource and capability based
theorizing (e.g. (Barney 1991; Teece, Pisano, and Shuen 1997)). However, the dominance of
the firm-level explanations have been partly refuted by later scholars finding the sum of
indirect and direct effects from industry to be of more importance than previously
indicated by (Mcgahan etc). Moreover, the industry effect is also more time invariant
(i.e. more stable) compared to firm level effects. Recent findings also suggest that there
are considerable interaction effects between industry and the firm indicating that
industry membership works directly at performance as well as indirectly through affecting
the firm. However, little or no explanations are made for the mechanisms of this
interaction.In other words, industry factors matter alas less than firm-level factors, and
the indirect link from industry-level through firms are in need of further development.

One way of looking at this is to consider the dynamic capabilities of the firm. Although
an increasing consensus focuses on these dynamic capabilities as the ability to create,
extend and modify firm resources in the face of change, there still looms an important
stream of literature suggesting that dynamic capabilities can materialize as management
actions (ref eisenhardt, peteraf, helfat). A management action aimed at strategic change
will very often entail "resource orchestration" as broadly discussed in the literature
(Helfat etc). But, strategic change are also found to come about from adapting and
changing the positioning of the firm within its industry (ref) or simply refocusing its
focal industry (ref). In other words, when a firm with excellent dynamic capabiliites
senses and seizes opportunities within its domain, the implementation of strategic change
will often both entail a considerable share of "asset orchestration" (helfat etc) as well
as management actions directed at positioning within its industry (ref). One way to
illustrate this is to consider Microsofts strategic change from operating system to cloud
provider. CEO Satya Nadela and his team was able to restart the business by creating,
extending and modifying the resources of the firm, including new skills, routiens and
hardware, as a response to a growing demand for cloud computing (ref nadellas
book). Meanwhile, his vision and formulated strategy also direclty put the company into a
new industry, or at the very least reshaped the position they were already in. While
Microsofts dynamic capabiliites worked significantly through "resource orchestration",
the sensing and seizing of the management also ordered a direct re-positioning at the same
time.

Moreover, The link between firm resources and competitive advantage is determined by the
extent to which resources are valuable and difficult to imitate \citep{Barney1991a}. This
principle is also used to evaluate how higher order capabilities (e.g. dynamic
capabilities \citep{Winter2003,Collis1994,Teece2007}. Hence, for a direct link to exist
the dynamic capabilities of the firm will have to be valuable and difficult to
imitate. This is an area of certain contention in the literature \citep{Peteraf2013} which
we get back to below, but an increasing majority is arguing that dynamic capabilities
reflect more complex processes that co-evolve with responses to market changes over time
and hence creating idiosyncratic capabilities that are unique, valuable and hard to copy
\citep{Helfat2007,Arndt2018}. Consequently, they are likely to lead to competitive
advantage.

In most cases such direct effects between dynamic capabilities and strategic change
resulting in competitive advantage, are combined with the indirect effects (i.e. resource
orchestration). To test this proposition empirically is hard, but one hypothesis following
from this is related to the role of mediators. if a direct path does not exist, the
relationship between dynamic capabilities and competitive advantage would be fully
mediated by the proper set of firm resources.

\emph{H1: The existence of a direct path from dynamic capabilities to competitive advantage
suggest that full mediation is not present}

Even if the dynamic capabilities of the firm themselves are not valuable and inimitable as
needed to create competitive advantage, the underlying resources of the firm may very well
be. Such resources in the form of operating routines or capabilities can in turn stem from
an inherent learning process evolutionary distinct to the firm. Operating routines can
thus lead to competitive advantage because of their evolutionary character
\citep{Nelson1982,Winter2003}. To the extent operating routines stems from idiosyncratic
evolutionary process of individual firms rather than being the result of best practices,
they can satisfy the VRIO condition and thus be a source of competitive advantage.  Hence,
through "resource orchestration" a firm with non-valuable dynamic capabilities can achieve
competitive advantage through the indirect path. The flip side of this is that a valuable
dynamic capability would not be able to create competitive advantage by orchestrating
non-valuable resources. Consequently, the mediation effects of underlying resources is
dependent on the nature of the resource - it being valuable and inimitable or not.

\emph{H2a: When dynamic capabilities are being mediated by VRIO resources the firm can
  achieve competitive advantage}

\emph{H2a: When dynamic capabilities are being mediated by VRIO resources the firm can not
  achieve competitive advantage}

DEspite a growing consensus on the nature of dynamic capabilites, a serious contention is
presen tin the literature \citep{Peteraf2013}. This stems from different notions of
dynamic capabilities in the two "founding" papers of the theory, \cite{Eisenhardt2000} and
\cite{Teece1997}.

One strand of contributions sees DC as routines that are simple in nature and more as best
practices to consider \citep{Eisenhardt2000}. Thus, dynamic capabilities are more like
behavioral patterns and less idiosyncratic to the firm and will consequently not be able
to satisfy the VRIO condition for competitive advantage to emerge. Dynamic capabilities as
such routines are short lived, unstable and “substitutable” (ibid, p 1110) thus violating
a key VRIO condition, and they are more “more homogeneous (..) than is usually assumed”
(ibid, 1116). However, the simplicity of simple rules also makes them faster do deploy and
put in motion, albeit potentially unstable. To the extent dynamic capabilities manifest
themselves as simple rules they cannot alone create competitive advantage, but could very
well generate strategic change by facilitating the process of resource orchestration.. In
other words, even if dynamic capabilities play out as simple rules that are not in
themselves satisfying the VRIO condition, the way they work through creating, extending
and modifying firm resources, can still yield competitive advantage through orchestrating
VRIO resources.

On the other hand, \cite{Teece1997} and others \cite{Helfat2007} argue that dynamic
capabilities will rather invariably be VRIO due to the evolutionary nature of their
development \citep{Arndt2018,Peteraf2013}. Arguing that dynamic capabilities reflect more
complex processes that co-evolve with responses to market changes over time and hence
creating idiosyncratic capabilities that are unique, valuable and hard to copy
\citep{Helfat2007,Arndt2018}. Consequently, they are likely to satisfy the VRIO condition
and thus leading to competitive advantage.

The contention between these streams of literature is most explicit in highly dynamic
environments (distefano, peteraf) suggesting that environmental dynamism could play into
the presence of direct and indirect linkages between dynamic capabilities and competitive
advantage. This is also supported in previous empirical literature \citep{Schilke2014}.
This could suggest that dynamic capabilities exhibit different characteristics under
various levels of environmental dynamism. This is also suggested by Eisenhardt blaboblabla
arguing that under moderatly dynamism dynamic capabilities takes the form of best
pracitces (e.g. aquisition targeting) and is less important than under more dynamic
environments. However, albeit best practices, the way they are implemented in the factor
markets are still idiosyncratic. And taken into conjunction with the routinzed process of
changing underlying routines, VRIO will very likely emerge under more stable
conditions. It is not that the market/direct expression of dynamic capabilities are not
present in stable environments, rather it is less pronounced and the indirect mediated
effect (evolutionary path) is more important. In other words, it is not neither nor, but
degrees of importance.

\emph{H3: The mediation effects of underlying resources is dependent on the environmental
dynamism facing the firm }

%% Increasingly the literature remain focused on an \emph{indirect
%%   path} between dynamic capabilities and competitive advantage
%% \citep{Barreto2010,Schilke2018}. The theory behind this path goes back to the core of
%% dynamic capabilities where resource orchestration happens as a response to changes in the
%% environment. When markets, technologies or other external factors changes, firms with high
%% levels of dynamic capabilities are able to seize opportunities for changes, seize these
%% opportunities and transform the resource base to fit the strategic change needed to stay
%% competitive \cite{Teece2007}. This ability to adapt is by some coined \emph{evolutionary
%%   fitness} defined as BLABLBALBA and stems from the ability of the firm to reconfigure
%% resources in the face of change. The extant literature, however, is less clear on how the
%% nature of the underlying resources matter in this indirect path. On the one hand we can
%% see

%% Following REF suggesting more reseaarch on dynamic capabilities should put forth
%% competeing hypotheses we propose these first

%% We introduce the notion of multilevel VRIO to integrate the interactions between different
%% types of capabilities at different levels. This entails that dynamic capabilities (the
%% high level) can consist of complex and evolutionary characteristics that are idiosyncratic
%% to the firm and hence VRIO by definition along the lines of how the concept is understood
%% by \cite{Teece1997}. On the other hand dynamic capabilities can also take the shape of
%% simpler routines which are short lived, unstable and substitutable thus violating a key
%% VRIO condition. This follows the notion of dynamic capabilities found in
%% h\cite{Eisenhardt2000} arguing that they are more “more homogeneous (..) than is usually
%% assumed” (EM: 1116).

%% Moreover, these dynamic capabilities are in the extant literature argued and found to work
%% directly \citep{Kor2005,Eriksson2014,Schilke2018} on competitive advantage which suggests
%% that they are indeed VRIO. On the other hand, a growing body of literature (ref) as well
%% a consensus on the definition of dynamic capabilities (ref), holds that they can lead to
%% competitive advantage by changing underlying resources and routines. In this
%% \emph{indirect path} the dynamic capabilities itself need not be VRIO as long as the
%% underlying resources themselves are.

%% \subsection{The role of dynamism}

%% The most contentious issue in the current debate on dynamic capabilities is how their
%% nature varies with the environment.

%% If htey are indeed latent

%% The lower level resources and routines are to be understood as conventional resources in
%% line with the resources-based view

%% WHAT IS THE ROLE OF DIRECT VS INDIRECT, HOW DOES IT RELATE TO THE ENVIRONMENT, AND WHAT
%% ROLE DOES THE TYPE OF UNDERLYING RESOURCE PLAY

%% Conventional dc literature has argued for both indirect and direct links. Indeed both
%% have been found, but while the indirect link is well rooted in the theory, the direct link
%% is to our knowledge less clear. The core idea of DC is that it works to modify underlying
%% resources as a response to changes in the environment and hence creating strategic change
%% needed to stay competitive. This resource orchestration is heavily represented in the
%% literature (ref) and through this process firms are able to generate CA. In such instances
%% the CA can stem from VRIO resources and/or VRIO DC. A VRIO DC would mean that the firm is able
%% to shuffle and reorganize generic resources in a way that creates a new and unique
%% configuration of resoures idiosyncratic and valuable to the firm. In such instances, the
%% outcome whould be CA. On the other hand, the DC may not by itself be VRIO but rather a set
%% of simple routines put in motion to change, but if the resources they reconfigure are
%% themselves VRIO, the outcome could just as eqsily be a competetive advantage. this
%% equfinaity is discussed in the literaure \cite{DiStefano2010} and makes it hard to
%% distingusih between the role of DC and ordinary capabilites in generating CA.

%% Furtemore, Considering a direct link between DC and CA, DC it self must consist of some
%% VRIO characteristics to garner CA. So in owther words, when is DC working directly and
%% indirectly_

%% How are dynamic capabilities working directly on competitive advantage? The microfoundations of dynamic capabilities explicated by \cite{Teece2007} is a useful
%% point of departure of this theoretical inquiry. He hold that for "analytical purposes,
%% dynamic capabilities can be disaggregated into the capacity (1) to sense and shape
%% opportunities and threats, (2) to seize opportunities, and (3) to maintain competitiveness
%% through enhancing, combining, protecting, and, when necessary, reconfiguring the business
%% enterprise’s intangible and tangible assets". Sensing and seizing is a part of what we
%% would coin the "opportunity capture" of dynamic capabiliites, where as reconfiguring
%% concerns itself with the "resource orchestration" needed to change assets to stay
%% competetive.

%% Hence, when considering indirect and direct paths to compettetive advantage we will have
%% to be clear on two fronts: Are DC themselves VRIO and are the underlying resources and
%% routines central to the resource orchestration, VRIO.

%% >

%% The microfoundations of dynamic capabilities—the distinct skills, processes, procedures, orga- nizational structures, decision rules, and disciplines—which undergird enterprise-level sensing, seizing, and reconfiguring capacities are difficult to develop and deploy.
%% For analytical purposes, dynamic capabilities can be disaggregated into the capacity (1) to sense and shape opportunities and threats, (2) to seize opportunities, and (3) to maintain competitiveness through enhancing, combining, protecting, and, when necessary, reconfiguring the business enter- prise’s intangible and tangible assets.

%%        name Mean_est     t
%% 3   DC -> P    0.396 6.157
%% 4   DC -> R    0.416 8.104
%% 5 DC -> CA1    0.239 4.018
%% 8  P -> CA1    0.172 2.724
%% 9  R -> CA1    0.081 1.262
%% > R
%%        Path   LOW     t  HIGH     t
%% 1 DC -> CA1 0.417 6.585 0.144 1.513
%% 2   DC -> R 0.388 6.078 0.534 7.525
%% 3  R -> CA1 0.017 0.219 0.211 2.123
%% > P
%%        Path   LOW     t  HIGH     t
%% 1 DC -> CA1 0.355 4.805 0.133 1.379
%% 2   DC -> P 0.386 5.294 0.436 4.663
%% 3  P -> CA1 0.155 2.193 0.232 2.315
%% > B
%%        name R_LOW_est   t.x R_HIGH_est   t.y
%% 1 DC -> CA1     0.355 5.407      0.044 0.400
%% 2   DC -> P     0.376 3.729      0.432 4.668
%% 3   DC -> R     0.366 5.569      0.522 7.225
%% 4  P -> CA1     0.158 1.989      0.220 2.357
%% 5  R -> CA1     0.010 0.135      0.200 2.033
%% >

%% %% creating and One example
%% %% is the role of post-acquisition routines where routines to identify and deal with
%% %% acquisition partners are a source of generating strategic change (Eisenhardt and Sull
%% %% 2001, ref). Another example is (see refs from Schilke 2018). The functioning of this
%% %% expression of dynamic capabilities are more latent and hence mainly observable ex
%% %% post. Hence we can see this as direct route to competitive advantage as this expression of
%% %% dynamic capabilities work directly in the factor market in capturing resources
%% %% needed for strategic change. We coin this direct connection as the \emph{market path}.

%% On the other hand, a stream of contributions see dynamic capabilities as evolutionary and
%% more complex routines (ref) or processes (ref) set to change underlying capabilities in
%% the face of changing conditions (TPS). This notion stems from evolutionary economics
%% \citep{Winter2003,Nelson1982} and the behavioral theory of the firm \citep{Cyert1963}
%% holding that path dependency and complex processes for changing underlying operating
%% routines are a source of competitive advantage for the firm. Along this indirect path
%% change is gradual and running through changes in underlying routines. We coin this the
%% \emph{evolutionary path}.

%% These paths from dynamic capabilities to competitive advantage are different expressions
%% of dynamic capabilities with different mediums of intermediation. The \emph{market path}
%% works in the market and his hence not typically altering lower order capabilities and
%% resources in the firm directly, but rather acquiring new ones. The source of competitive
%% advantage along this path stems from the interaction with the other, \emph{evolutionary
%%   path} in a dynamic bundle \citep{Peteraf2013}. Such interactions create a bundle of
%% capabilities and routines that are socially complex and hard to imitate
%% \citep[p. 320]{DiStefano2014}. Moreover, the interactions between the two paths of dynamic
%% capabilities is separable with respect to what intermediation exist. Consequently,
%% separate paths exist albeit a part of a dynamic bundle.

%% \cite{DiStefano2014} propose an integrated framework where simple routines and more
%% complex processes could contribute simultanously to the firms outcome, specifically its
%% competitive advantage. In this framework, coined the "organizational drivetrain" the
%% authors envision the movement of an organization as set in motion by a drivetrain system
%% not unlike that of a bicycle.

%% Building on the notion of an organizational drivetrain we argue that dynamic capabilities
%% can indeed impact the organizational movement or change through two distinct but
%% interconnected mechanisms (i.e. the crank set and the rear shifter). The two mechanisms
%% consitutting the drivetrain is consistent with our notion of \emph{evolutionary path} and the \emph{market path}. The
%% evolutionary path is instigated by more complex processes setting in motiong a chance in
%% the lower order (operational) routines \cite{Collis1994,Winter2003}. In the drivetrain
%% metaphor this path corresponds to change stemming from the front crank set of thich there
%% are fewer gears, but with larger impact. This path is slower
%% and yields a more idiosyncratic set of organizational routines which is very much in line
%% with the idea put forth in the seminal founding papers \citep{Teece1997,Winter2003}.

%% Seeing the bike rider and her efforts to handle the levers of control (i.e. the gear
%% shifter) as the dynamic capability, and consider it not only as a management effort, but
%% that of the organization as a whole. The process of changing front gears are more profound
%% (i.e. evolutionary) and costly, and it is for larger and slower strategic
%% changes. Operating with simple routines in factor markets, however, is quicker and the
%% rider does this on a more frequent basis with less effort.

%% H1: Dynamic capabilities works to create competitive advantage through two distinctly
%% different paths.
